---
title: "Introduction to luzlogr"
author: "Ben Bond-Lamberty"
date: "October 4, 2015"
output: html_document
---

This vignette describes how to use the **luzlogr** package, which provides flexible but lightweight logging facilities for R scripts.

## Installing
To install this package:

1. Make sure you have `devtools` installed from CRAN and loaded.
2. `install_github("bpbond/luzlogr")`

## Basics
Three functions - `openlog()`, `printlog()`, `closelog()` - provide logging of script output. They're simple to use:

```{r}
library(luzlogr)
openlog("test.log")
printlog("message")
closelog()
```

The resulting log file `test.log` looks like this (not including the initial `##` characters):
```{r, echo=FALSE}
lg <- readLines("test.log")
invisible(file.remove("test.log"))
for(i in seq_along(lg)) cat(lg[i], "\n")
```

Any printable object can be written to a log. For example:

```{r}
openlog("test.log")
printlog("message", 1, 2)
printlog(head(cars))
closelog(sessionInfo = FALSE)
```

(Notice we've told `closelog()` not to append `sessionInfo()` output, as it does by default.)

```{r, echo=FALSE}
lg <- readLines("test.log")
invisible(file.remove("test.log"))
for(i in seq_along(lg)) cat(lg[i], "\n")
```

## More involved examples

By design, `luzlogr` is intended to be flexible and easy to use. Nonetheless, it does provide additional features, including:

* priority levels for logs and messages
* capturing *all* script output
* logging to a text file or arbitrary [connection](https://stat.ethz.ch/R-manual/R-devel/library/base/html/connections.html)

### Priority levels

Need - solution - example

### Capturing all output

Need - solution - example

### Logging to a connection

Need - solution - example

This concludes the **Introduction** vignette.
