---
title: "Introduction to luzlogr"
author: "Ben Bond-Lamberty"
date: "October 4, 2015"
output: html_document
---

This vignette describes how to use the **luzlogr** package, which provides flexible but lightweight logging facilities for R scripts.

## Installing
To install this package, use R's `install.packages()` function, or the "Packages" pane in [RStudio](https://www.rstudio.com/products/RStudio/). To install the latest version of this package from GitHub:

1. Make sure you have the `devtools` package installed from CRAN and loaded.
2. `install_github("bpbond/luzlogr")`

## Basics
Three functions - `openlog()`, `printlog()`, `closelog()` - provide logging of script output. They're simple to use:

```{r}
library(luzlogr)
openlog("test.log")
printlog("message")
closelog()
```

The resulting log file `test.log` looks like this (not including the initial `##` characters):
```{r, echo=FALSE}
lg <- readLines("test.log")
invisible(file.remove("test.log"))
for(i in seq_along(lg)) cat(lg[i], "\n")
```

By default individual lines are prefixed with a timestamp, and end with a carriage return, and the entire log has starts with an open message and ends with a close one.

Any printable object can be written to a log. For example:

```{r}
openlog("test.log")
printlog("message", 1, 2)
printlog(head(cars))
closelog(sessionInfo = FALSE)
```

(Notice we've told `closelog()` not to append `sessionInfo()` output, as it does by default.)

```{r, echo=FALSE}
lg <- readLines("test.log")
invisible(file.remove("test.log"))
for(i in seq_along(lg)) cat(lg[i], "\n")
```

## More involved examples

By design, `luzlogr` is intended to be simple and easy to use. Nonetheless, it does provide additional features, including:

* priority levels for logs and messages
* flagging messages
* capturing *all* script output
* logging to a text file or arbitrary [connection](https://stat.ethz.ch/R-manual/R-devel/library/base/html/connections.html)
* switching logs

### Priority levels

In many circumstances, we want messages to have different priorities or *levels* associated with them. If the logfile itself has a minimum priority level, set when it's opened, this makes it easy to produce logs of varying levels of detail.

By default, **luzlogr** logs have a minimum levels of `-Inf`: in other words, they'll accept every single message sent to them via `printlog()`, which produces messages with a default level of zero. We can change the log's level, however, and this will then affect which messages appear in the log.

```{r}
openlog("test.log", loglevel = 0)
printlog("This message will appear", level = 0)
printlog("So will this (level 0 by default)")
printlog("This will not", level = -1)
closelog(sessionInfo = FALSE)
```

produces

```{r, echo=FALSE}
lg <- readLines("test.log")
invisible(file.remove("test.log"))
for(i in seq_along(lg)) cat(lg[i], "\n")
```

Notice that the third message didn't get logged. If we change the `loglevel` parameter in `openlog()` to -1 or lower, however, all these messages will appear.

### Flagging messages

Another way to differentiate messages is by *flagging* them. Note that in all the above examples, when the log is closed, a `flags = 0` message is printed. But we can change that:

```{r}
openlog("test.log")
printlog("A normal message")
printlog("A flagged message!", flag = TRUE)
flaglog("Another")
closelog(sessionInfo = FALSE)
```

```{r, echo=FALSE}
lg <- readLines("test.log")
invisible(file.remove("test.log"))
for(i in seq_along(lg)) cat(lg[i], "\n")
```

### Capturing all output

So far, only messages sent via `printlog()` appear in the log. We might, however, want to capture *everything*^[Almost. Messages, warnings, and errors will not appear.] produced by a script. To do this, use the `sink = TRUE` option of `openlog()`.

### Logging to a connection

Logs can also be sent to any R [connection](https://stat.ethz.ch/R-manual/R-devel/library/base/html/connections.html): a pipe, compressed file, URL, etc.

```{r}
con <- gzfile("test.log.gz")
openlog(con)
printlog("Sending to a compressed logfile")
closelog(sessionInfo = FALSE)
```

```{r, echo=FALSE}
invisible(file.remove("test.log.gz"))
```

Note that **luzlogr** won't close a connection that was already open at the time of the `openlog()` call.

### Switching logs

If you're logging to log A and open log B (without closing A), subsequent `printlog()` messages will go to B. When you close B, logging switches back to A (i.e., there's a stack of logs that gets pushed/popped when necessary).

If you need to append to an existing log, use `append = TRUE` when calling `openlog()`. By default, existing logfiles are erased upon opening.


This concludes the **Introduction** vignette.
